|

### 2023-09-22 JCGS_response_to_reviews/response_to_review.Rmd

---
title: "Response to Reviews: JCGS-23-139"
author: Weihao Li, Emi Tanaka, Dianne Cook, Susan VanderPlas
output: pdf_document
date: "`r Sys.Date()`"
bibliography: ../ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We thank the anonymous associate editor and reviewers for their careful reviews. What follows is our point-by-point response to reviewer comments.

@@

Note that *the original reviewer comments are in italic* and our response is in normal text.

Text in the paper corresponding to the changes as a response to reviewers' comments are highlighted in red. We have also made a lot of minor changes smoothing the language and fixing small issues in the paper which are not highlighted.

@@

*1. "The ubiquity of this advice is curious: investigating why, is the subject of this paper. " I don't find this advice curious - it seems almost universal. I have taught this topic for many years and give the same advice. I explain why I give this advice and I believe several other authors do the same. Their reasons are wider in scope than you express in this article.*

We have changed the language.

*2. All your examples are simulated data and very obviously so. Real data shows much greater irregularity - the points will usually not be uniformly spaced along the x-axis. The response may be somewhat discrete or have other features that will distinguish it from bland randomness. This would seem to offer greater challenges to your lineup idea as generating data under the null will be much more challenging. If we simulate from the null model, the real data will be very obvious. You have done an experiment but the reality would be more like an observational study. This limits the scope of your conclusions which you should clearly acknowledge.*

The work intentionally uses simulated data because we need to study the problem in controlled conditions. The reviewer is correct, though, that real data will pose more interesting challenges. It should be noted that our simulation setup does include discrete independent variables, and consequently somewhat discretized response. The other point to note is that our null generating process uses residual rotation and will likely maintain irrelevant quirks of the data in the null plots as well. In that sense, the real data plot is not necessarily always obvious (see @buja2009statistical and @wickham2010 for examples). Nonetheless, our simulation does only test handful of conditions. We have addressed this limited scope in the broader discussion of the use of residual plots and lineups for residual diagnostics added to the new Section "Limitations and practicality". <!-- (see issue #11).-->

@@

<!--
ET: I don't think talking about computer vision model makes sense here as the reviewer would not know about our other work?

An observational study involving actual residual plots will hopefully be possible once a computer vision model is deployed that will allow analysts to upload their residuals for automated testing.
-->

*3. It is true that formal diagnostic tests tend to be too sensitive and yet we don't have a definite sense of how much divergence from the assumed model would cause a significant problem. How much non-normality in the residuals is acceptable? We don't really know. Formal tests give us definite answers about violations of assumptions which we might find hard to ignore but diagnostic plots give us license to sweep inconvenient features under the carpet. Is there really an advantage to the diagnostic plot in this case other than the opportunity to hand-wave about our answer?*

When the diagnostic plots are embedded in a lineup of nulls, it does have the advantage of providing a calibrated answer, and protects against the subjectiveness of hand-waving. We are advocating that you should always calibrate the diagnostic plot by a comparison with null plots. We have strengthened this point of view by adding more explanation in the Introduction, in addition to what is provided in the Conclusions.

Having a definitive (numerical) answer to "how much is too much" would require more and different work. For our purposes, it was that the true residual plot could be detected from the null plots when the lineup was evaluated by people. We've added a sentence to the Conclusion clarifying this.

@@

<!--(Is there a paper to refer the reader to?)-->

*4. I use something like the lineup protocol in teaching to help students calibrate their assessment of diagnostic plots. But I doubt very much experienced analysts use them in practice. I do not think it is realistic or really that helpful to recommend their everyday use.*

It can be argued that it is being used by practitioners based on download rates of the `nullabor` package (>19,000 downloads in the past year) which implements a process for creating lineups, and application articles that report usage. Using it formally, we agree requires the effort of engaging a crowd-sourcing service, involving a small expense. Science labs typically have a budget for equipment and experimental devices and we would advocate that this is a useful addition. Where a practitioner lacks the budget for running crowd-sourcing services, the protocol can be used with their team or network and can be an excellent way to engage their team/network with the data. We have formally used crowd-sourcing ourselves for application purposes in @roychowdhuryUsingVisualStatistical2015a for ecological data and @mahbubulmajumderVisualMiningMethods2013 for gene expression analysis. We have added some details in the "Limitations and practicality" section. <!--We agree that it is not widespread, and think that this is because it's a little unwieldy. The goal is to make it easier, particularly by employing an automated testing tool using computer vision, but this is not the subject of the current paper.-->

@@

*This article uses the lineup protocol in a visual inference experiment, which puts a residual plot in the context of null plots. Focus is mostly on detecting nonlinearities and heteroskedasticity. Comparisons with some classic tests are given.*

*Overall Comments*

*Overall, this paper is very well written and interesting. However, I have some serious reservations about the practicality of the procedure, which I articulate in greater detail below. I also think the paper is currently too long, and that a significant chunk of material should be shifted to the Appendix.*

Thank you for the positive feedback. We have shortened the paper by about five pages, while also adding a discussion on the practicality in the Section "Limitations and practicality".

@@

Thank you for your suggestions. This is how we have shortened the paper:

- Details of Section 3.2 Statistical significance moved to the Appendix. 
- Sections 1 and 2 have been made more precise, while still covering the requested additional topics.
- The explanation of power calculation was left in the main paper because this is new and of primary importance for understanding the results.
- Part of the Experimental design has been moved into the Appendix.

*2. The practical implementation of this as a procedure is, in my opinion, the biggest weakness. The authors acknowledge the costliness of human evaluation of residuals, but not until Page 33, Line 47. I find this a tad disingenuous as this will likely jump out to the reader as being a problem right from the beginning of reading this paper (it did for me). So I think being upfront about this limitation, and providing some practical considerations to address this concern will assuage my concerns. There might be some concepts from human-computer interaction that could be effective for operationalizing and improving the efficiency of the present work.*

The new Section "Limitations and practicality" addresses these concerns and we have added a couple of sentences to the Introduction (highlighted in red).

### 2023-09-22 JCGS_response_to_reviews/response_to_review.Rmd

*This article uses the lineup protocol in a visual inference experiment, which puts a residual plot in the context of null plots. Focus is mostly on detecting nonlinearities and heteroskedasticity. Comparisons with some classic tests are given.*

*Overall Comments*

*Overall, this paper is very well written and interesting. However, I have some serious reservations about the practicality of the procedure, which I articulate in greater detail below. I also think the paper is currently too long, and that a significant chunk of material should be shifted to the Appendix.*

Thank you for the positive feedback. We have shortened the paper by about four pages, while also adding a discussion on the practicality in the Section "Limitations and practicality".

@@

This is not computationally expensive. The conventional test also make mistakes. You need something more than the conventional test. And you also need something more than the residual plot.

check earlier paper.-->

*4. What should be claimed about the sample size n as a limitation? Take in point the comments at the bottom of page 19. Am I to understand that the lineup protocol will not be appropriate for big data problems?*

The lineup protocol is very suitable for big data. A plot other than a scatterplot is probably needed to better display a large number of observations. We have changed the language so that this is clearer (highlighted in red).

### 2023-09-22 paper.Rmd
changes reviewed, and small fixes made

\textcolor{red}{However, relying purely on subjective assessment of a single plot can be deceiving. Indeed, correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task. People will almost always see a pattern \citep[see ][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise. \cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}

\textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot. For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern. In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot. When used this way, it provides an objective test for data plots. \cite{majumdervalidation2013} validated that results from lineups assessed by human observers performed similarly to conventional tests. One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct. However, where no adequate conventional test exists, it is invaluable, as shown by \cite{loy2013diagnostic}. Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}

@@

@@

```{r first-example-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot $2^2 + 2$, exhibiting non-linearity) is embedded among 19 null plots, where the residuals are simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot $2^2 + 2$."}

@@

One of the primary limitations of the lineup protocol lies in its reliance on human judgements. In this context, the effectiveness of a single lineup evaluation is closely connected with the perceptual ability and visual skills of the individual. When results from multiple individuals are combined the outcome is encouragingly robust. For simple plots and strong patterns just a few individuals are needed to arrive at a clear answer, but more individuals will be needed when the plot design is complex, or the signal strength is weak. <!--This reliance on human judgment introduces an element of variability and subjectivity that can potentially impact the robustness of the results obtained. To ensure reliable outcomes, it's often necessary to include a diverse group of participants capable of offering unbiased judgements. This requirement has led researchers, including those in our study, to turn to crowd-sourcing platforms that offer access to a random selection of participants, which can help mitigate potential bias.-->

@@

The experiment also revealed some interesting results. For the most part, the visual test performed similarly to the appropriate conventional test with a shift in the power curve. Unlike conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals. If the fitted value distribution is not uniform, there is a small loss of power in the visual test. Surprisingly there is a big difference in power of the conventional test across fitted value distributions. Another surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.

>>>>

\textcolor{red}{Additionally, relying solely on the subjective assessment of a single plot can be problematic. People will almost always see a pattern \citep[see][]{kahneman2011thinking}, so the question that really needs answering is whether any pattern perceived is consistent with randomness, or sampling variability, or noise. Correctly judging whether \emph{no} pattern exists in a residual plot is a difficult task. \cite{loy2021bringing} emphasizes that this is especially difficult to teach to new analysts and students, and advocates to the broader use of the lineup protocol \citep{bujastatistical2009}.}

\textcolor{red}{The lineup protocol places a data plot in a field of null plots, allowing for a comparison of patterns due purely by chance to what is perceived in the data plot. For residual analysis this is especially helpful for gauging whether there is \emph{no} pattern. (Figure \ref{fig:first-example-lineup} shows an example of a lineup of residual plots.) In its strict use, one would insist that the data plot is not seen before seeing the lineup, so that the observer doesn't know which is the true plot. When used this way, it provides an objective test for data plots. \cite{majumdervalidation2013} validated that results from lineups assessed by human observers performed similarly to conventional tests. One wouldn't use a lineup when a conventional test exists and is adequate because it is more manually expensive to conduct. However, where no adequate conventional test exists, it is invaluable, as shown by \cite{loy2013diagnostic}. Here we use the lineup as a vehicle to rigorously explore why experts advise that residual plots are indispensable despite the prevalence of numerical tests.}

@@

```{r first-example-lineup, fig.pos="t!", fig.height = 7, fig.width = 7, fig.cap = "Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data is embedded among 19 null plots, where the residuals are simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting the data plot (at position $2^2 + 2$, exhibiting non-linearity)."}

@@

@@

One of the primary limitations of the lineup protocol lies in its reliance on human judgements. In this context, the effectiveness of a single lineup evaluation can be dependent on the perceptual ability and visual skills of the individual. However, when results from multiple individuals are combined the outcome is encouragingly high-quality and robust. For simple plots and strong patterns just a few individuals are needed to arrive at a clear answer, but more individuals will be needed when the plot design is complex, or the signal strength is weak. <!--This reliance on human judgment introduces an element of variability and subjectivity that can potentially impact the robustness of the results obtained. To ensure reliable outcomes, it's often necessary to include a diverse group of participants capable of offering unbiased judgements. This requirement has led researchers, including those in our study, to turn to crowd-sourcing platforms that offer access to a random selection of participants, which can help mitigate potential bias.-->

@@

The experiment also revealed some interesting results. For the most part, the visual test performed similarly to the appropriate conventional test with a shift in the power curve. Unlike conventional tests, where one needs to specifically test for non-linearity or heteroskedasticity the visual test operated effectively across the range of departures from good residuals. If the fitted value distribution is not uniform, there is a small loss of power in the visual test. Surprisingly, there is a big difference in power of the conventional test across fitted value distributions. Another unexpected finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it: both triangles being more difficult to detect than the butterfly, and a small difference in detection between left- and right-triangle.
